An Algorithm for Searching Evolving Graphs in Linear Algebraic Language

### 2016-01-21 paper.tex
update introduction

Let's imagine a game played by three people $A$, $B$ and $C$.
Suppose we give a piece of information $X$ to $A$ and ask 
$A$ talks to $B$ first without $C$'s presence and then ask $B$ talks to $C$ without $A$'s presence.
In a different day, we give a piece of information $Y$ to $A$ again.
But this time, we ask $B$ talks to $C$ first and 
then ask $A$ talks to $B$ without $C$'s presence. 
We now ask the question ``Will $C$ find out what $X$ is ? Will $C$ find out
what $Y$ is?''

Evolving graphs are models to study this kind of scenario where the 
interrelations between entities are time-dependent.
% We assume that the reader is familiar with the basic graph terminology. 
% All the necessary material can be found in}.
We follow the notation of (static) graphs in  \cite{even12} and \cite{kegi11}.
In general, evolving graphs $G(t) = (V(t), E(t))$ are graphs that change as a function of time $t$. However, in different papers the term ``evolving graph'' has been
used to refer to slightly different models. 
 Flajolet et al. \cite{fkp89} use the term to mean the evolution of random graphs.
Bahmani et al. \cite{bkmu12} use the term to mean a sequence of graphs where
the difference between two consecutive graphs is small.
Similar models are also investigated using the term ``evolving network'' and 
``time-dependent graph''.

In this work, we consider a discretization of the timeline as $n$ distinct 
timestamps. We define an evolving graph $G_n$ to be 
a sequence of (static) graphs 
$G_n = \langle G^{[1]}, G^{[2]}, \ldots, G^{[n]} \rangle$. 
Depends on applications, the graph $G^{[i]}$ could mean either
a snapshot of $G(t)$ at timestamp $i$ or  the union of all the graphs
between timestamp $i-1$ and $i$ (excludes timestamp $i-1$).
Both the node set and edge set of $G^{[i]}$ could be different from
these of $G^{[j]}, j \ne i$. 
\begin{center}
\begin{tikzpicture}[snake=zigzag, line before snake = 5mm, line after snake = 5mm]
    % draw horizontal line   
    \draw (0,0) -- (2,0);
    \draw[snake] (2,0) -- (4,0);
    \draw (4,0) -- (5,0);
    \draw[snake] (5,0) -- (7,0);

@@

% why BFS is important? 
Breadth-first search is probably the most basic graph searching algorithm and 
it is also the building block of other graph algorithms like minimum spanning tree 
and shortest path \cite{ckrs09}.
Given a graph and a source node $s$, 
the breadth-first search (BFS) systematically discover every node that is reachable from $s$. 
As the name suggested, BFS expands the frontier of discovered regions
uniformly across the breadth, i.e., the algorithm discovers all nodes
that are $k$ walks away from $s$ before discovering any nodes that are
$k+1$ walks away.

A temporal path is the basic element to 
 define almost  all important evolving graph algorithms.
Matrix-matrix multiplication between consecutive adjacency matrices
were commonly used as a means of determining temporal paths \cite{gphe11} \cite{grihig13}. However, we show
in one specific case it fails to include all the temporal paths and thus should not be used 
for searching evolving graphs in general.
Despite the fact that graph metrics \cite{ntmm13} \cite{tmml09} \cite{tsmm09} and 
various centrality measures (including Katz  \cite{gphe11} \cite{grihig13}, 
betweenness \cite{alhi15}, and closeness  \cite{tmml10}) 
have been generalized for evolving graphs,
a BFS for evolving graphs, which we shall call it \emph{evolving-graph BFS},
was never explicitly described or verified.

@@

\section{Node Activeness}
\label{sec:matr-matr-mult}

@@

@@

\section{Application in Citation Networks}
\label{sec:applications}

>>>>

Let's imagine a game played by three people
$A$, $B$, and $C$. Suppose each of them has a piece of information,
say $X$, $Y$, and $Z$ respectively. The goal of the game is 
to get information from other people as much as possible. 
Suppose $A$ talks to $B$ first and then $B$ talks to $C$, 
then $C$ may has all three pieces of information $X$, $Y$, and $Z$ even 
if $C$ has never talked to $A$. 
However, if $B$ talks to $C$ first and then $A$ talks to $B$, then
no matter how smart $C$ is, $C$ can never get information 
$X$.  Time is an important factor for information flow.

We can analyze the spread of information between $A$, $B$, and $C$
as a process on a graph that changes over time.  
Such graphs are known as evolving graphs.
% Evolving graphs are models to study this kind of scenario where the 
% interrelations between entities are time-dependent.
% We assume that the reader is familiar with the basic graph terminology. 
% All the necessary material can be found in}.
We follow the notation of (static) graphs in  \cite{even12} and \cite{kegi11}.
In general, a evolving graph $G(t) = (V(t), E(t))$ is  a graph where the node set
and edge set are functions of time $t$.
However, in different literatures the term ``evolving graph'' has been
used to refer to slightly different models. 
 Flajolet et al. \cite{fkp89} use the term to mean the evolution of random graphs.
Bahmani et al. \cite{bkmu12} use the term to mean a sequence of graphs where
the difference between two consecutive graphs is small.
Similar models are also investigated using the term ``evolving network'' and 
``time-dependent graph''.

In this work, we consider a discretization of the timeline as $n$ distinct 
timestamps and define an evolving graph $G_n$ to be 
a sequence of (static) graphs 
$G_n = \langle G^{[1]}, G^{[2]}, \ldots, G^{[n]} \rangle$. 
Depends on applications, the graph $G^{[i]}$ could mean either
a snapshot of $G(t)$ at timestamp $i$ or  the union of all the graphs
between timestamp $i-1$ and $i$ (excludes timestamp $i-1$).
Both the node set and edge set of $G^{[i]}$ could be different from
these of $G^{[j]}, j \ne i$. 
\begin{center}
\begin{tikzpicture}[snake=zigzag, line before snake = 5mm, line after snake = 5mm]
    % draw horizontal line   
    \draw (0,0) -- (2,0);
    \draw[snake] (2,0) -- (4,0);
    \draw (4,0) -- (5,0);
    \draw[snake] (5,0) -- (7,0);

@@

% why BFS is important?

% Given a graph and a source node $s$, 
% the breadth-first search (BFS) systematically discover every node that is reachable from $s$. 
% As the name suggested, BFS expands the frontier of discovered regions
% uniformly across the breadth, i.e., the algorithm discovers all nodes
% that are $k$ walks away from $s$ before discovering any nodes that are
% $k+1$ walks away. 
A temporal path is the basic element to 
 define almost  all important evolving graph algorithms.
Matrix-matrix multiplication between consecutive adjacency matrices
were commonly used as a means of determining temporal paths \cite{gphe11} \cite{grihig13}. However, we show
in one specific case it fails to include all the temporal paths and thus should not be used 
for searching evolving graphs in general.

Breadth-first search is probably the most basic graph searching algorithm and 
it is also the building block of other graph algorithms like minimum spanning tree 
and shortest path \cite{ckrs09}.
Despite the fact that graph metrics \cite{ntmm13} \cite{tmml09} \cite{tsmm09} and 
various centrality measures (including Katz  \cite{gphe11} \cite{grihig13}, 
betweenness \cite{alhi15}, and closeness  \cite{tmml10}) 
have been generalized for evolving graphs,
a BFS for evolving graphs, which we shall call it \emph{evolving-graph BFS},
was never explicitly described or verified.

@@

\section{Node Activeness and its Impact}
\label{sec:matr-matr-mult}

\subsection{Aggregated Graphs}
\label{sec:aggregated-graphs}

In this section, we answer the question why we need evolving graph.

At the beginning of the paper, we set up a game played by three people $A$, $B$,
and $C$ and we showed if

@@

Introducing ones on the diagonal of the matrices will still get the wrong result.

@@

\section{Application to Citation Networks}
\label{sec:applications}
